---
title: "NIH Vector Biology Submission"
author: "sbsambado"
date: "8/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# pacakages
library(readr)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(ggpubr)
library(gridExtra)

```


Original Excel sheet is called longformat_questingticks08292020.csv
```{r dataframes}
#### make my separate dataframes ####

# full dataframe
# updated 09092020 with Susan's data
long <- read.csv("~/Desktop/Proposals:Grants/NIH 2020 Swei, Briggs, Sambado/figures/Excel sheets for figures/longformat_questingticks.csv")

# miyamotoi dataframe
miyamotoi <- filter(long, borrelia == 'miyamotoi')

# burgdorferi dataframe
burgdorferi <- filter(long, borrelia == 'burgdorferi')

# miyamotoi & nymphs
bmN <- filter(long, borrelia == 'miyamotoi' & lifestage == 'nymph')

# miyamotoi & adults
bmA <- filter(long, borrelia == 'miyamotoi' & lifestage == 'adult')

# burgdorferi & nymphs
bbN <- filter(long, borrelia == 'burgdorferi' & lifestage == 'nymph')

# burgdorferi & adults
bbA <- filter(long, borrelia == 'burgdorferi' & lifestage == 'adult')


```


Figure 1 compares prevalence of B. burgdroferi (pink) and B. miyamotoi (blue) at individual sites within a region. The points are broken into nymphs (triangles) and adults (circles), with the label of total ticks sampled at that site. For the northeastern sites, it's very clear that B. burgdorferi occurs at higher levels than B. miyamotoi (which we already knew & Rick's unpublished data confirms). This pattern holds up in the midwest at varying degrees, but I would like to incorporate Susan's data (if possible) to distinguish between northwestern/southeastern midwest sites. For the westcoast, there is no clear pattern besides that B. burgdorferi and B. miyamotoi fluctuate site by site. 

```{r Fig1_prevalence_by_sitesID_region.jpeg}

### base ggplot

## this is chunky as one plot, so I'll separate it into 3 graphs and plot onto one page later in this chunk, but good base start
p <- ggplot(data = long, # full dataframe
            aes(x = site_ID, y = prev, # prev ~ site_ID
                group = borrelia, color = borrelia)) + # group by BM or BB
  facet_wrap(~region, nrow = 3, scales = 'free_y') + # separate graph by region, scale to give unique y-axis values/region
  geom_line() +
  geom_point()+
  geom_text(aes(label = total_ticks, hjust = 1.8, vjust = 0), size = 1.4) + # add total tick count next to individual point
  xlab('Individual Sites') + ylab('Prevalence of Questing Ticks') +
theme(panel.border = element_blank(), # white background
        panel.grid.major = element_blank(), # white background
        panel.grid.minor = element_blank(), # white background
        axis.line = element_line(color = 'black')) +
  theme_classic()  # clean background 
p + theme(axis.text.x = element_text(angle = 45, size = 7)) # angle site labels at 45 degrees, not sure why this argument didn't work in previous chunk

#################### ^^^^ I would like to fix legend, point text labels, the labels for facet groups ###############


######## less chunky graph, split ggplots into 3 separate regions and then paste onto one page with ggpubr() ######

## first filter dramaframes into separate regions
NE <- filter(long, region == 'northeast')
MW <- filter(long, region == 'midwest')
WC <- filter(long, region == 'westcoast')


# Northeastern plot
n <-ggplot(data = NE, aes(x = site_ID, y = prev, group = borrelia, color = borrelia))+
  geom_line()+
  geom_point(aes(shape = lifestage))+
  geom_text(aes(label = total_ticks, hjust = 1.8, vjust = 0), size = 2) +
  xlab('Individual Sites') + ylab('Prevalence of Questing Ticks') +
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = 'black'),
        axis.text.x = element_text(angle = 45)) +
  theme_classic()+
  ggtitle('Northeastern US')

### full plot for northeastern
ne <- n + theme(axis.text.x = element_text(angle = 90, size = 7),
                axis.title = element_text(face = 'bold'),
                plot.title = element_text(face = 'bold', hjust = 0.5)) 

# Midwestern plot
m <-ggplot(data = MW, aes(x = site_ID, y = prev, group = borrelia, color = borrelia))+
  geom_line()+
  geom_point(aes(shape = lifestage)) +
  geom_text(aes(label = total_ticks, hjust = 1.8, vjust = 0), size = 2) +
  xlab('Individual Sites') + ylab('Prevalence of Questing Ticks') +
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = 'black'),
        axis.text.x = element_text(angle = 45)) +
  theme_classic()+
  ggtitle('Midwestern US')


### full plot for midwest
mw <- m + theme(axis.text.x = element_text(angle = 90, size = 7),
                axis.title = element_text(face = 'bold'),
                plot.title = element_text(face = 'bold', hjust = 0.5)) 


# Western plot

w <-ggplot(data = WC, aes(x = site_ID, y = prev, group = borrelia, color = borrelia))+
  geom_line()+
  geom_point(aes(shape = lifestage)) +
  geom_text(aes(label = total_ticks, hjust = 1.8, vjust = 0), size = 2) +
  xlab('Individual Sites') + ylab('Prevalence of Questing Ticks') +
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = 'black'),
        axis.text.x = element_text(angle = 45)) +
  theme_classic() +
  ggtitle('Western US')

wc <- w + theme(axis.text.x = element_text(angle = 90, size = 7),
                axis.title = element_text(face = 'bold'),
                plot.title = element_text(face = 'bold', hjust = 0.5)) 

############# FINAL PRODUCT FOR FIG 1 ######################

full <-ggarrange(ne + rremove('legend') + rremove('xlab'),  #ggarrange is how your image will look like when it's printed 
          mw + rremove('legend') + rremove('ylab') + rremove('xlab'), #rremove removes items attached to that string 
          wc + rremove('ylab') + rremove('xlab'),
          ncol = 3, common.legend = TRUE, legend = 'bottom',
          align = "hv") # only 1 legend printed at the bottom
annotate_figure(full,
                bottom = text_grob("Individual Sites", face = "bold"))

## ^^ would like to fix legend and make text points more visable 
```

just collaborators data

```{r}
NE <- filter(long, author == 'barbour')
MW <- filter(long, author == 'paskewitz')
WC <- filter(long, author == 'sambado')


# Northeastern plot
n <-ggplot(data = NE, aes(x = site_ID, y = prev, group = borrelia, color = borrelia))+
  geom_line()+
  geom_point(aes(shape = lifestage))+
  geom_text(aes(label = total_ticks, hjust = 1.8, vjust = 0), size = 2) +
  xlab('Individual Sites') + ylab('Prevalence of Questing Ticks') +
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = 'black'),
        axis.text.x = element_text(angle = 45)) +
  theme_classic()+
  ggtitle('Northeastern US')

### full plot for northeastern
ne <- n + theme(axis.text.x = element_text(angle = 90, size = 7),
                axis.title = element_text(face = 'bold'),
                plot.title = element_text(face = 'bold', hjust = 0.5)) 

# Midwestern plot
m <-ggplot(data = MW, aes(x = site_ID, y = prev, group = borrelia, color = borrelia))+
  geom_line()+
  geom_point(aes(shape = lifestage)) +
  geom_text(aes(label = total_ticks, hjust = 1.8, vjust = 0), size = 2) +
  xlab('Individual Sites') + ylab('Prevalence of Questing Ticks') +
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = 'black'),
        axis.text.x = element_text(angle = 45)) +
  theme_classic()+
  ggtitle('Midwestern US')


### full plot for midwest
mw <- m + theme(axis.text.x = element_text(angle = 90, size = 7),
                axis.title = element_text(face = 'bold'),
                plot.title = element_text(face = 'bold', hjust = 0.5)) 


# Western plot

w <-ggplot(data = WC, aes(x = site_ID, y = prev, group = borrelia, color = borrelia))+
  geom_line()+
  geom_point(aes(shape = lifestage)) +
  geom_text(aes(label = total_ticks, hjust = 1.8, vjust = 0), size = 2) +
  xlab('Individual Sites') + ylab('Prevalence of Questing Ticks') +
  theme(panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = 'black'),
        axis.text.x = element_text(angle = 45)) +
  theme_classic() +
  ggtitle('Western US')

wc <- w + theme(axis.text.x = element_text(angle = 90, size = 7),
                axis.title = element_text(face = 'bold'),
                plot.title = element_text(face = 'bold', hjust = 0.5)) 

############# FINAL PRODUCT FOR FIG 1 ######################

full <-ggarrange(ne + rremove('legend') + rremove('xlab'),  #ggarrange is how your image will look like when it's printed 
          mw + rremove('legend') + rremove('ylab') + rremove('xlab'), #rremove removes items attached to that string 
          wc + rremove('ylab') + rremove('xlab'),
          ncol = 3, common.legend = TRUE, legend = 'bottom',
          align = "hv") # only 1 legend printed at the bottom
annotate_figure(full,
                bottom = text_grob("Individual Sites", face = "bold"))
```




Figure 2 is less detailed, but I wanted to visualize Rick's idea to compare the min/max prevalence of each borrelia in their respective regions. For me, the main takeaway is that B. miyamotoi has greater variation between east and west coast, but the median prevalence of B. miyamotoi remains consistent, so maybe we end back where we left off.
```{r Fig2_minmax_prevalence_by_region.jpeg}

# finding the min/max of BM (also median)
bmminmax <-ggplot(data = miyamotoi) + # pull from just bm dataset
  stat_summary( # stat_summary can do a bunch of fun things!
    mapping = aes(x = region, y = prev),
    fun.min = min, 
    fun.max = max,
    fun = median) + # find median
  theme_classic() +
  xlab('Regions') + ylab('Prevalence of Questing Ticks') +
  ggtitle('Range of B. miyamotoi (Min/Max)')+ 
  theme(axis.text.x = element_text(face = 'bold', size = 12), # improve font
                axis.title = element_text(face = 'bold'),
                plot.title = element_text(face = 'bold', hjust = 0.5)) 

# finding the min/max of BB (also median)
bbminmax<-ggplot(data = burgdorferi) +
  stat_summary(
    mapping = aes(x = region, y = prev),
    fun.min = min, 
    fun.max = max,
    fun = median) +
  theme_classic() +
  xlab('Regions') + ylab('Prevalence of Questing Ticks') +
  ggtitle('Range of B. burgdorferi (Min/Max)')+ 
  theme(axis.text.x = element_text(face = 'bold', size = 12),
                axis.title = element_text(face = 'bold'),
                plot.title = element_text(face = 'bold', hjust = 0.5)) 

  
### FINAL PRODUCT with combined graphs
minmax <- ggarrange(bmminmax, bbminmax+ rremove('ylab'), # ggarrange fixes your plot in it's final output form
          ncol = 2)

#### ^^^ would like to fix the order (summary_stat is blocking my arguments), title, try to combine on 1 graph
```

```{r extra graphs}
## full dataset (prev ~ site) faceted by state
ggplot(data = long) +
  geom_point(mapping = aes(x = site_ID, y = prev, color = borrelia)) +
  facet_wrap(~ state, nrow = 2)

## full dataset (prev ~ site) faceted by state with jitter
ggplot(data = long) +
  geom_point(mapping = aes(x = site_ID, y = prev, color = borrelia), position = 'jitter') + facet_wrap(~ region, nrow = 3, scale = 'free_y') +
  theme_bw()

## full dataset (prev ~ region) 
ggplot(data = long) +
  geom_point(mapping = aes(x = region, y = prev, color = borrelia))

## miyamotoi dataset (prev ~ siteID) by lifestage
ggplot(data = miyamotoi) +
  geom_point(mapping = aes(x = site_ID, y = prev, color = lifestage))

## coordflip to visualize FULL data easier with siteID on vertical, could color coordinate this by latitude
ggplot(data = long, mapping = aes(x = site, y = prev)) +
  geom_boxplot() +
  coord_flip()

## coordflip to visualize BM data easier with siteID on vertical, could color coordinate this by latitude
ggplot(data = miyamotoi, mapping = aes(x = state, y = prev)) +
  geom_boxplot() +
  coord_flip()

## coordflip to visualize BB data easier with siteID on vertical, could color coordinate this by latitude
ggplot(data = burgdorferi, mapping = aes(x = state, y = prev)) +
  geom_boxplot() +
  coord_flip()



```

```{r code that needs to be tweaked more}

# add CI 95%, need to fix code
sts <- summarySE(long, measurevar = 'prev', groupvars = c('burgdorferi', 'miyamotoi'))
ggplot(long, aes(x = site_ID, y = prev, color = borrelia))+
    geom_errorbar(aes(ymin = prev - se, ymax = len + se), width = .1) +
    geom_line()+
    geom_point()

```

